Simple repression draft

### 2020-05-29 doc/section_02_means.tex
Many additions to main text Bayesian inference

\subsubsection{Summary of Results Ccross models for Fold-Changes in Mean
Expression}
What have we learned from our exploration of the models
in~\fig{fig1:means_cartoons} and their predictions of fold-changes in mean
expression levels? The key outcome of our analysis is the existence of a master
curve shown in \fig{fig1:means_cartoons}(D) to which the fold-change predictions
of all the models collapse. This master curve is parametrized by only two
effective parameters: $\Delta F_R$, which characterizes the number of repressors
and their binding strength to the DNA, and $\rho$, which characterizes all other
features of the promoter architecture. The key assumption underpinning this
result is that no transcription occurs when a repressor is bound to its
operator. Note, however, that we are agnostic about the molecular mechanism
which achieves this; steric effects are one plausible mechanism, but, for
instance, ``action-at-a-distance'' mediated by kinked DNA due to repressors
bound tens or hundreds of nucleotides upstream of a promoter is plausible as
well.

@@

Moving forward, we have weak evidence supporting the interpretation of $\Delta
F_R$ as a physically real free energy~\cite{Landman2019} and other work casting
doubt~\cite{Hammar2014}. How might we resolve the confusion? If there is no
discriminatory power to test the theory and distinguish the various models with
measurements of fold-changes in means, how do we probe the theory? Clearly to
discriminate between the nonequilibrium models in~\fig{fig1:means_cartoons}, we
need to go beyond means to ask questions about kinetics, noise and even full
distributions of mRNA copy numbers over a population of cells. If the
``one-curve-to-rule-them-all'' is more than a mathematical tautology, then the
free energy of repressor binding inferred from fold-change measurements should
agree with repressor binding and unbinding rates. In other words, the
equilibrium and nonequilibrium definitions of $\Delta F_R$ must agree, meaning
\begin{equation}
\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-),
\end{equation}
must hold, where $\beta\Delta\varepsilon_R$ is inferred from the master curve
fit to fold-change measurements, and $k_R^+$ and $k_R^-$ are inferred in some
orthogonal manner. Single molecule measurements such as from~\cite{Hammar2014}
have directly observed these rates, and in the remainder of this work we explore
a complementary approach: inferring repressor binding and unbinding rates
$k_R^+$ and $k_R^-$ from FISH measurements of mRNA population distributions.
\marginpar{\it MM: Add a notation aside somewhere that all association rates are written as zeroth-order rates. In other words, the concentration of the molecule is hidden inside, so for instance, $k_R^+$ is proportional to repressor copy number.}

>>>>

\subsubsection{Summary of Results across models for Fold-Changes in Mean
Expression}
What have we learned from our exploration of the models
in~\fig{fig1:means_cartoons} and their predictions of fold-changes in mean
expression levels? The key outcome of our analysis is the existence of a master
curve shown in \fig{fig1:means_cartoons}(D) to which the fold-change predictions
of all the models collapse. This master curve is parametrized by only two
effective parameters: $\Delta F_R$, which characterizes the number of repressors
and their binding strength to the DNA, and $\rho$, which characterizes all other
features of the promoter architecture. The key assumption underpinning this
result is that no transcription occurs when a repressor is bound to its
operator. Note, however, that we are agnostic about the molecular mechanism
which achieves this; steric effects are one plausible mechanism, but, for
instance, ``action-at-a-distance'' mediated by kinked DNA due to repressors
bound tens or hundreds of nucleotides upstream of a promoter is plausible as
well.

@@

Moving forward, we have weak evidence supporting the interpretation of $\Delta
F_R$ as a physically real free energy~\cite{Landman2019} and other work casting
doubt~\cite{Hammar2014}. How might we resolve the confusion? If there is no
discriminatory power to test the theory and distinguish the various models with
measurements of fold-changes in means, how do we probe the theory? Clearly to
discriminate between the nonequilibrium models in~\fig{fig1:means_cartoons}, we
need to go beyond means to ask questions about kinetics, noise and even full
distributions of mRNA copy numbers over a population of cells. If the
``one-curve-to-rule-them-all'' is more than a mathematical tautology, then the
free energy of repressor binding inferred from fold-change measurements should
agree with repressor binding and unbinding rates. In other words, the
equilibrium and nonequilibrium definitions of $\Delta F_R$ must agree, meaning
\begin{equation}
\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-),
\label{eq:deltaFR_eq_noneq_equiv}
\end{equation}
must hold, where $\beta\Delta\varepsilon_R$ is inferred from the master curve
fit to fold-change measurements, and $k_R^+$ and $k_R^-$ are inferred in some
orthogonal manner. Single molecule measurements such as from~\cite{Hammar2014}
have directly observed these rates, and in the remainder of this work we explore
a complementary approach: inferring repressor binding and unbinding rates
$k_R^+$ and $k_R^-$ from FISH measurements of mRNA population distributions.
\marginpar{\it MM: Add a notation aside somewhere that all association rates are written as zeroth-order rates. In other words, the concentration of the molecule is hidden inside, so for instance, $k_R^+$ is proportional to repressor copy number.}

### 2020-05-29 doc/section_03_beyond_means.tex
Many additions to main text Bayesian inference

In this section, our objective is to explore the same models considered in the
previous section, but now with reference to the the question of how well they
describe the distribution of gene expression levels, with special reference to
the variance in these distributions. To that end, we repeat the same pattern as
in the previous section by examining the models one by one. \mrm{In particular
we will focus on the Fano factor, defined as the variance/mean. This metric
serves as a discriminatory tool from the null model that the steady-state mRNA
distribution must be Poisson, giving a Fano factor of one.}

>>>>

In this section, our objective is to explore the same models considered in the
previous section, but now with reference to the the question of how well they
describe the distribution of gene expression levels, with special reference to
the variance in these distributions. To that end, we repeat the same pattern as
in the previous section by examining the models one by one. In particular
we will focus on the Fano factor, defined as the variance/mean. This metric
serves as a powerful discriminatory tool from the null model that the
steady-state mRNA distribution must be Poisson, giving a Fano factor of one.

### 2020-05-29 doc/section_04_bayesian_inference.tex
Many additions to main text Bayesian inference

\textcolor{red}{In this section of the paper, we continue our program of providing
one complete description of the entire broad sweep of studies
that have been made in the context of the repressor-operator model,
dating all the way back to the original work of Jacob and Monod
and including the visionary quantitative work of M\"{u}ller-Hill and collaborators
and up to more recent studies.   In addition, the aim is to reconcile the 
equilibrium and non-equilibrium perspectives that have been brought
to bear on this problem.  To that end, we require a self-consistent means
of determining the parameters in these different theoretical analyses and
the current section does that, first by examining the constitutive promoter
and then turning to the full case of simple repression. }

@@

From consideration of Fano factors in the previous section, we suspect that
model 5 in Figure~\ref{fig2:constit_cartoons}(A), a one-state bursty model of
constitutive promoters, achieves a Goldilocks level of complexity. Does this
stand up to closer scrutiny, namely, comparison to full mRNA distributions
rather than simply their moments? We will test this thoroughly on
single-cell mRNA counts for different unregulated promoters from Jones et.\
al.~\cite{Jones2014}.

@@

It is interesting to connect this observation to the work
of~\cite{Brewster2012}, where these same 18 promoters were considered through
the lens of the three-state equilibrium model (model 2 in
Figure~\ref{fig1:means_cartoons}(B)) and binding energies $\Delta\varepsilon_P$
were predicted from an energy matrix model derived from~\cite{Kinney2010}. As
previously discussed the thermodynamic models of gene regulation can only make
statements about the mean gene expression. This implies that we can draw the
connection between both frameworks by equating the mean mRNA $\left\langle m
\right\rangle$. This results in
\begin{equation}
\langle m \rangle = \frac{k_i b}{\gamma}
        = \frac{r}{\gamma}
        \frac{\frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P)}
                {1+\frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P)}.
\end{equation}
By taking the weak promoter approximation for the equilibrium model ($P/N_{NS} 
\exp(-\beta\Delta\varepsilon_r) \ll 1$) results in
\begin{equation}
\langle m \rangle = \frac{k_i b}{\gamma}
        = \frac{r}{\gamma} \frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P),
\end{equation}
valid for all the binding energies considered here.

@@

\subsection{Transcription factor kinetics can be inferred from FISH
measurements}\label{sec:rep_kinetics_inference}

Now that we have a satisfactory model in hand for constitutive promoters, we
would like to return to the main thread: can we infer repressor binding and
unbinding rates from mRNA distributions over a population of cells as measured
by single-molecule FISH in~\cite{Jones2014}? If so, how do these inferred rates
compare to direct single-molecule measurements such as from~\cite{Hammar2014}
and to binding energies such as from~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}, which were inferred under the assumptions of the
equilibrium models in Figure~\ref{fig1:means_cartoons}(B)? And can this
comparison shed light on the unreasonable effectiveness of the equilibrium
models?

@@

As we found in Section~\ref{sec:beyond_means}, for our purposes the ``right''
model of a constitutive promoter is the bursty picture, model five in
Figure~\ref{fig2:constit_cartoons}. Therefore our starting point here is the
analogous model with repressor added, model 5 in
Figure~\ref{fig1:means_cartoons}(C). For a given repressor binding site and copy
number, this model has four rate parameters to be inferred: the repressor
binding and unbinding rates $k_R^+$, and $k_R^-$, the initiation rate of bursts,
$k_i$, and the mean burst size $b$ (we nondimensionalize all of these by the
mRNA degradation rate $\gamma$).

Before showing the mathematical formulation of our statistical inference model,
we would like to sketch the intuitive structure. The dataset
from~\cite{Jones2014} we consider consists of single-cell mRNA FISH measurements
of nine different conditions, spanning several combinations of three unique
repressor binding sites and four unique repressor copy numbers. We assume that
the values of $k_i$ and $b$ are known, since we have already cleanly inferred
them from constitutive promoter data, and further we assume that these values
are the same across datasets with different repressor binding sites and copy
numbers. In other words, we assume that the regulation of the transcription
factor does not affect the mean burst size nor the burst initiation rate. The
regulation happens as the promoter is taken away from the transcriptionally
active state when the promoter is bound. We assume that there is one unbinding
rate parameter for each repressor binding site, which is shared across aTc
inducer conditions, and likewise one binding rate for each unique repressor copy
number. This makes our model nine dimensional, or nine if one counts $k_i$ and
$b$ as well.

Formally now, denote the set of seven repressor rates to be inferred as
\begin{equation}
\vect{k} =\{k_{Oid}^-, k_{O1}^-, k_{O2}^-,
k_{0.5}^+, k_{1}^+, k_{2}^+, k_{10}^+\},
\end{equation}
where subscripts for dissociation rates $k^-$  indicate the different repressor
binding sites, and subscripts for association rates $k^+$ indicate the aTc
concentration at which the repressor itself was induced. We emphasize that for
this particular dataset the repressor copy numbers are not known directly, so we
label their association rates by the concentration of a small molecule inducer
aTc which indirectly controls the transcription of $\textit{lacI}$. Also note
that the authors of~\cite{Jones2014} report estimates of LacI copy number per
cell rather than direct measurements. However, these estimates were made
\textit{assuming} the validity of the equilibrium models in
Figure~\ref{fig1:means_cartoons}, and since \textit{testing} these models is our
goal, we will make no assumptions about the LacI copy number for given aTc
concentrations.

@@

We ran MCMC sampling on the full nine dimensional posterior specified by this
generative model. To attempt to visualize this object, in
Figure~\ref{fig4:repressed_post_full}(A) we plot several two-dimensional slices
as contour plots, analogous to Figure~\ref{fig:constit_post_full}(C). Each of
these nine slices corresponds to the $(k_R^+, k_R^-)$ pair of rates for one of
the conditions from the dataset used to fit the model and gives a
sense of the uncertainty and correlations in the posterior.
\mrm{before invoking part (C) of the figure we are missing th explanation for 
part (B). I don't fully understand how these computations were made, so I'll 
let Jupiter detail this.}

>>>>

In this section of the paper, we continue our program of providing
one complete description of the entire broad sweep of studies
that have been made in the context of the repressor-operator model,
dating all the way back to the original work of Jacob and Monod
and including the visionary quantitative work of M\"{u}ller-Hill and collaborators
and up to more recent studies.   In addition, the aim is to reconcile the 
equilibrium and non-equilibrium perspectives that have been brought
to bear on this problem. From Section~\ref{section_02_means}, this reconciliation
depends on a key quantitative question as codified by Eq.~\ref{eq:deltaFR_eq_noneq_equiv}: does the free energy of repressor binding, as described in the equilibrium models and indirectly inferred from gene expression measurements, agree with the corresponding values of repressor binding and unbinding rates in the non-equilibrium picture, measured or inferred more directly? In this section we tackle the statistical inference problem of inferring these repressor rates from mRNA FISH data. But before we can turn to the full case of simple repression, we must choose an appropriate model of the constitutive promoter and infer the parameter values in that model. This is the problem we address first.

@@

From consideration of Fano factors in the previous section, we suspect that
model 5 in Figure~\ref{fig2:constit_cartoons}(A), a one-state bursty model of
constitutive promoters, achieves the right balance of complexity
and simplicity, by allowing both Fano factor $\nu>1$, but also by
remedying, by design, the problems of parameter degeneracy than
model 4 in Figure~\ref{fig2:constit_cartoons} suffered. Does this
stand up to closer scrutiny, namely, comparison to full mRNA distributions
rather than simply their moments? We will test this thoroughly on
single-cell mRNA counts for different unregulated promoters from Jones et.\
al.~\cite{Jones2014}.

@@

\subsubsection{Connecting inferred parameters to prior work}
It is interesting to connect these speculations on $k_i$ and $b$ to the work
of~\cite{Brewster2012}, where these same 18 promoters were considered through
the lens of the three-state equilibrium model (model 2 in
Figure~\ref{fig1:means_cartoons}(B)) and binding energies $\Delta\varepsilon_P$
were predicted from an energy matrix model derived from~\cite{Kinney2010}. As
previously discussed the thermodynamic models of gene regulation can only make
statements about the mean gene expression. This implies that we can draw the
connection between both frameworks by equating the mean mRNA $\left\langle m
\right\rangle$. This results in
\begin{equation}
\langle m \rangle = \frac{k_i b}{\gamma}
        = \frac{r}{\gamma}
        \frac{\frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P)}
                {1+\frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P)}.
\end{equation}
By taking the weak promoter approximation for the equilibrium model ($P/N_{NS} 
\exp(-\beta\Delta\varepsilon_r) \ll 1$) results in
\begin{equation}
\langle m \rangle = \frac{k_i b}{\gamma}
        = \frac{r}{\gamma} \frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P),
\end{equation}
valid for all the binding energies considered here.

@@

\subsection{Transcription factor kinetics can be inferred from FISH
measurements}\label{sec:rep_kinetics_inference}
\subsubsection{Building the model and performing parameter inference}

Now that we have a satisfactory model in hand for constitutive promoters, we
would like to return to the main thread: can we reconcile the equilibrium
and nonequilibrium models by putting to the test
Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, the correspondence between indirectly
inferred equilibrium binding energies and nonequilibrium kinetic rates?
To make this comparison, is it possible to infer repressor binding and
unbinding rates from mRNA distributions over a population of cells as measured
by single-molecule FISH in~\cite{Jones2014}? If so, how do these inferred rates
compare to direct single-molecule measurements such as from~\cite{Hammar2014}
and to binding energies such as from~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}, which were inferred under the assumptions of the
equilibrium models in Figure~\ref{fig1:means_cartoons}(B)? And can this
comparison shed light on the unreasonable effectiveness of the equilibrium
models, for instance, in their application in~\cite{Chure2019, Chure2019a}?

@@

As we found in Section~\ref{sec:beyond_means}, for our purposes the ``right''
model of a constitutive promoter is the bursty picture, model five in
Figure~\ref{fig2:constit_cartoons}A. Therefore our starting point here is the
analogous model with repressor added, model 5 in
Figure~\ref{fig1:means_cartoons}(C). For a given repressor binding site and copy
number, this model has four rate parameters to be inferred: the repressor
binding and unbinding rates $k_R^+$, and $k_R^-$, the initiation rate of bursts,
$k_i$, and the mean burst size $b$ (we nondimensionalize all of these by the
mRNA degradation rate $\gamma$).

Before showing the mathematical formulation of our statistical inference model,
we would like to sketch the intuitive structure. The dataset
from~\cite{Jones2014} we consider consists of single-cell mRNA FISH measurements
of nine different conditions, spanning several combinations of three unique
repressor binding sites and four unique repressor copy numbers. We assume that
the values of $k_i$ and $b$ are known, since we have already cleanly inferred
them from constitutive promoter data, and further we assume that these values
are the same across datasets with different repressor binding sites and copy
numbers. In other words, we assume that the regulation of the transcription
factor does not affect the mean burst size nor the burst initiation rate. The
regulation happens as the promoter is taken away from the transcriptionally
active state when the promoter is bound by repressor. We assume that there is one unbinding
rate parameter for each repressor binding site, which is shared across aTc
inducer conditions, and likewise one binding rate for each unique repressor copy
number. This makes our model seven dimensional, or nine if one counts $k_i$ and
$b$ as well. Note that we use only a subset of the datasets from
Jones et.\ al.~\cite{Jones2014}, as discussed more in Appendix~\ref{sec:bayes}.

Formally now, denote the set of seven repressor rates to be inferred as
\begin{equation}
\vect{k} =\{k_{Oid}^-, k_{O1}^-, k_{O2}^-,
k_{0.5}^+, k_{1}^+, k_{2}^+, k_{10}^+\},
\end{equation}
where subscripts for dissociation rates $k^-$  indicate the different repressor
binding sites, and subscripts for association rates $k^+$ indicate the aTc
concentration at which the repressor itself was induced. We emphasize that for
this particular dataset the repressor copy numbers are not known directly, so we
label their association rates by the concentration of a small molecule inducer
aTc which indirectly controls the transcription of $\textit{lacI}$, as
described in detail in~\cite{Jones2014}. Also note
that the authors of~\cite{Jones2014} report estimates of LacI copy number per
cell rather than direct measurements. However, these estimates were made
\textit{assuming} the validity of the equilibrium models in
Figure~\ref{fig1:means_cartoons}, and since \textit{testing} these models is our
present goal, it would be circular logic if we were to make the same assumption.
Therefore we will make no assumptions about the LacI copy number for given aTc
concentrations.

@@

We ran MCMC sampling on the full nine dimensional posterior specified by this
model. To attempt to visualize this object, in
Figure~\ref{fig4:repressed_post_full}(A) we plot several two-dimensional slices
as contour plots, analogous to Figure~\ref{fig:constit_post_full}(C). Each of
these nine slices corresponds to the $(k_R^+, k_R^-)$ pair of rates for one of
the conditions from the dataset used to fit the model and gives a
sense of the uncertainty and correlations in the posterior.
We note that the 95\% uncertainties of all the rates span about $\sim0.3$
log units, or about a factor of two, with the exception of $k_{0.5}^+$, the
association rate for the lowest repressor copy number which is somewhat larger.

\mmnote{These next 2 paragraphs should be SI, I think. Leaving here for
now to avoid a merge conflict w/ MRM.}

We found that fitting a single operator/aTc concentration at a
time with a single binding and unbinding rate did not yield a
stable inference for most of the possible operator/aTc
combinations. In other words, a single dataset could not
independently resolve the binding and unbinding rates, only their
ratio as set by the mean fold-change in
Figure~\ref{fig1:means_cartoons}. Only by making the assumption
of a single unique binding rate for each repressor copy number
and a single unique unbinding rate for each binding site, as done
in Figure~\ref{fig4:repressed_post_full}A, was it possible to
independently resolve the rates and not merely their ratios.

We also note that we found it necessary to exclude the very
weakly and very strongly repressed datasets from Jones et.\
al.~\cite{Jones2014}. In both cases there was, in a sense, not
enough information in the distriubtions for our inference
algorithm to extract, and their inclusion simply caused problems
for the MCMC sampler without yielding any new insight. For the
strongly repressed data (Oid, 10~ng/mL aTc), with >95\% of cells
with zero mRNA, there was quite literally very little data from
which to infer rates. And the weakly repressed data, all with the
repressor binding site O3, had an unbinding rate so fast that the
sampler essentially sampled from the prior; the likelihood had
negligible influence, meaning the data was not informing the
sampler in any meaninful way, so no inference was possible.

\subsubsection{Comparison with prior measurements of repressor binding energies}
Our primary goal in this work is to reconcile the kinetic
and equilibrium pictures of simple repression.
Towards this end we would like to compare the repressor kinetic rates we
have inferred with the repressor binding energies inferred through
multiple methods in~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}.
If the agreement is close, then it suggests that the equilibrium
models are not wrong and the repressor binding energies they contain
correspond to physically real free energies, not mere fit parameters.

Figure~\ref{fig4:repressed_post_full}(B) shows both comparisons,
with the top panel comparing to equilbrium binding energies
and the bottom panel comparing to single-molecule measurements.

First consider the top panel and its comparison between repressor
kinetic rates and binding energies.
As described in section~\ref{section_02_means}, if the
equilibrium binding energies from~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018} indeed are the physically real binding
energies we believe them to be, then they should be related to
the repressor kinetic rates via
Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, which we reprint here,
\begin{equation}
\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-).
\label{eq:deltaFR_eq_noneq_equiv_repeat}
\end{equation}
Assuming mass action
kinetics implies that $k_R^+$ is proportional to repressor copy
number $R$, or more precisely, it can be thought of as repressor
copy number times some intrinsic per molecule association rate.
But since $R$ is not directly known for our data
from~\cite{Jones2014}, we cannot use this equation directly.
Instead we can consider two different repressor binding sites and
compute the \textit{difference} in binding energy between them,
since this difference depends only on the unbinding rates and not
on the binding rates. This can be seen by evaluating
Eq.~\ref{eq:deltaFR_eq_noneq_equiv_repeat} for two different
repressor binding sites, labeled (1) and (2), but with the same
repressor copy number $R$, and taking the difference to find
\begin{equation}
\Delta F_R^{(1)} - \Delta F_R^{(2)}
= \beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= - \log(k_R^+/k_1^-) + \log(k_R^+/k_2^-),
\end{equation}
or simply
\begin{equation}
\beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= \log(k_2^-/k_1^-).
\end{equation}
The left and right hand sides of this equation are exactly the horizontal
and vertical axes of the top panel of Figure~\ref{fig4:repressed_post_full}.
Since we inferred rates for three repressor binding sites (O1, O2, and Oid),
there are only two independent differences that can be constructed, and we
arbitrarily chose to plot O2-O1 and O1-Oid in
Figure~\ref{fig4:repressed_post_full}(B).
Numerically, we compute values of $k_{O1}^- / k_{Oid}^-$ and
$k_{O2}^- / k_{O1}^-$ directly from our full posterior samples,
which conveniently provides uncertainties as well, as detailed in
Appendix~\ref{sec:bayesian}.
\mmnote{Explain in appendix how marginalization with sampling makes this
``trivial'' to compute from samples of the joint posterior, b/c
sampling makes marginalization trivial.}
We then compare these ratios of rates to the binding energy differences
$\Delta\epsilon_{O1} - \Delta\epsilon_{Oid}$ and from
$\Delta\epsilon_{O2} - \Delta\epsilon_{O1}$ as computed from the values
from both~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}.
Three of the four values are within $\sim0.5 k_BT$ of the diagonal
representing perfect agreement, which is comparable to the
$\sim\text{few}\times0.1 k_BT$ variability between the independent
determinations of the same quantities between~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}. The only outlier involves Oid measurements
from~\cite{Razo-Mejia2018}, and as the authors of~\cite{Razo-Mejia2018} note,
this is a difficult measurement of low fluorescence signal against
high background since Oid represses so strongly. We are therefore
inclined to regard the failure of this point to fall near the
diagonal as a testament to the difficulty of the measurement and
not as a failure of our theory.
\mmnote{Does this seem a fair characterization?}

On the whole then, we regard this as striking confirmation of the
validity of the equilibrium models. Their lynchpin parameter is a
phenomenological free energy of repressor binding that has
previously only been inferred indirectly. Our result shows that
the microscopic interpretation of this free energy, as the log of
a ratio of transition rates, does indeed hold true.
\mmnote{Does this deserve qualifiers, like ``to a good approximation,'' or ``as a first order model'' or something? I like the simple finality without, but I worry it's too brazen or subtlely misleading.}

\subsubsection{Comparison with prior measurements of repressor kinetics}
In the previous section we established the equivalence between
the equilibrium models' binding energies and the repressor
kinetics we infer from mRNA population distributions.
But one might worry that the repressor rates we infer
from mRNA distributions are \textit{themselves} merely
fitting parameters and that they do not actually correspond to the
binding and unbinding rates of the repressor in vivo.
To verify that this is not the case, we next compare our kinetic rates with a
different measurement of the same rates using a radically different method:
single molecule measurements as performed in Hammar et.\ al.~\cite{Hammar2014}.
This is plotted in the lower panel of Figure~\ref{fig4:repressed_post_full}B.

Since we do not have access to repressor copy number for either
the FISH data from~\cite{Jones2014} or the single-molecule data
from~\cite{Hammar2014}, we cannot make an apples-to-apples
comparison of association rates $k_R^+$. Further, while Hammar
et.\ al.\ directly measure the dissociation rates $k_R^-$, our
inference procedure returns $k_R^-/\gamma$, i.e., the repressor
dissociation rate nondimensionalized by the mRNA degradation rate
$\gamma$. So to make the comparison, we must make an assumption
for the value of $\gamma$ since it was not directly measured. For
most mRNAs in \textit{E.\ coli}, quoted values for the typical mRNA
lifetime $\gamma^{-1}$ range between about 2.5~min~\cite{Chen2015}
to 8~min. We chose $\gamma^{-1} = 3$~min and $\gamma^{-1} =
5$~min as representative values and plot a comparison of
$k_{O1}^-$ and $k_{Oid}^-$ from our inference with corresponding values
reported in~\cite{Hammar2014} for both these choices of $\gamma$.

The degree of quantitative agreement in the lower panel of
Figure~\ref{fig4:repressed_post_full}B clearly depends on the
precise choice of $\gamma$. Nevertheless we find this comparison very satisfying, when two wildly different approaches to a measurement of the same quantity yield broadly compatible results. We emphasize the agreement between our rates
and the rates reported in~\cite{Hammar2014} for any reasonable
$\gamma$: values differ by at most a factor of 2 and possibly
agree to within our uncertainties of 10-20\%. From this we feel
confident asserting that the parameters we have inferred from
Jones et.\ al.'s FISH data do in fact correspond to repressor
binding and unbinding rates, and therefore our conclusions on the
agreement of these rates with binding energies
from~\cite{Garcia2011a} and~\cite{Razo-Mejia2018} are valid.

\subsubsection{Model checking}
In Figure~\ref{fig:constit_post_full}B we saw that the simple
Poisson model of a constitutive promoter, despite having a well
behaved posterior, was clearly insufficient to describe the data.
It behooves us to carry out a similar check for our model of
simple repression, codified by Eq.~\ref{eq:p_m_bursty+rep} for
the steady-state mRNA copy number distribution.
As derived in Sections~\ref{section_02_means}
and~\ref{sec:beyond_means}, we have compelling theoretical reasons
to believe it is a good model, but if it nevertheless turned out
to be badly contradicted by the data we should like to know.

The details are deferred to Appendix~\ref{sec:bayesian}, and here
we only attempt to summarize the intuitive ideas, as detailed at
greater length by Jaynes~\cite{Jaynes2003} as well as Gelman and
coauthors~\cite{Gelman2013,Gelman2013a}. From our samples of the
posterior distribution, plotted in
Figure~\ref{fig4:repressed_post_full}A, we generate many
replicate data using a random number generator.
In Figure~\ref{fig4:repressed_post_full}C, we plot ECDFs of the
middle 95\% quantiles of these replicate data with the actual
experimental data from Jones et.\ al.~\cite{Jones2014} overlaid,
covering all ten experimental conditions spanning repressor
binding sites and copy numbers (as well as the constitutive
baseline UV5).

The purpose of Figure~\ref{fig4:repressed_post_full}C is simply a
graphical, qualitative assessment of the model: do the
experimental data systematically disagree with the simulated
data, which would suggest that our model is missing important
features? A further question is not just whether there is a
detectable difference between simulated and experimental data,
but whether this difference is likely to materially affect the
conclusions we draw from the posterior in
Figure~\ref{fig4:repressed_post_full}A. More rigorous and
quantitative statistical tests are possible~\cite{Gelman2013},
but their quantitativeness does not necessarily make them more
useful. As stated in~\cite{Gelman2013a}, we often find this
graphical comparison more enlightening because it better engages
our intuitition for the model, not merely telling \textit{if} the
model is wrong but suggesting \textit{how} the model may be
incomplete.

Our broad brush takeaway from
Figure~\ref{fig4:repressed_post_full}C is overall of good
agreement. There some oddities, in particular the long tails in
the data for Oid, 1~ng/mL, and O2, 0.5~ng/mL. The latter is
especially odd since it extends beyond the tail of the
unregulated UV5 distribution. This is a relatively small number
of cells, however, so whether this is a peculiarity of the
experimental data, a statistical fluke of small numbers, or a
real biological effect is unclear. It is conceivable that there
is some very slow timescale switching dynamics that could cause
this bimodality, although it is unclear why it would only appear
for specific repressor copy numbers. There is also a small offset
between experiment and simulation for O2 at the higher repressor
copy numbers, especially at 2~ng/mL. We suspect this may be
caused by the very large number of zero mRNA counts for Oid,
2~ng/mL skewing it's partner datasets through the shared
association rate, and so we are not too concerned by this
discrepancy either. None of these fairly minor quibbles cause us
to seriously doubt the overall correctness of our model, which
further validates its use to compare the equilbrium models'
binding energies to the nonequilibrium models' repressor
kinetics, as we originally set out to do.
